\pdftitle

### 2018-05-18 maxNt_v3.tex
V3

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-18T10:23:24+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

%\newcommand*{\vs}{{vs.}}

@@

\section{Maximum-entropy: sample level vs full-population level}
\label{sec:specific_initial_probability}

@@

The difference between the applications at the sample and at the population
levels appears also in statistical mechanics. If a system is statistically
described by a \me\ Gibbs state, its subsystems cannot be described by a
Gibbs state \cite{maesetal1999}. A somewhat similar situation also appears
in the statistical description of the final state of a non-equilibrium
process starting and ending in two equilibrium states: we can describe our
knowledge about the final state either by (1) a Gibbs distribution,
calculated from the final equilibrium macrovariables, or (2) by the
distribution obtained from the Liouville evolution of the Gibbs
distribution assigned to the initial state. The two distributions differ
(even though the final \emph{physical} state is obviously exactly the same
\cite[\sect~4]{jaynes1985d_r1993}), and the second allows us to make
sharper predictions about the final physical state thanks to our knowledge
of its preceding dynamics. In this example, though, both distributions are
usually immensely sharp and practically lead to the same predictions. In
the neuroscientific applications considered in this note the difference in
predictions of the sample vs full-population applications may be very
relevant instead.

>>>>

%% Author: PGL  Porta Mana
%% Created: 2015-11-04T09:20:08+0100
%% Last-Updated: 2018-05-18T10:25:17+0200
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Report-no: ***
\documentclass{article}

@@

\newcommand*{\vs}{{vs}}

@@

\section{Maximum-entropy: sample level \vs\ full-population level}
\label{sec:specific_initial_probability}

@@

The difference between the applications at the sample and at the population
levels appears also in statistical mechanics. If a system is statistically
described by a \me\ Gibbs state, its subsystems cannot be described by a
Gibbs state \cite{maesetal1999}. A somewhat similar situation also appears
in the statistical description of the final state of a non-equilibrium
process starting and ending in two equilibrium states: we can describe our
knowledge about the final state either by (1) a Gibbs distribution,
calculated from the final equilibrium macrovariables, or (2) by the
distribution obtained from the Liouville evolution of the Gibbs
distribution assigned to the initial state. The two distributions differ
(even though the final \emph{physical} state is obviously exactly the same
\cite[\sect~4]{jaynes1985d_r1993}), and the second allows us to make
sharper predictions about the final physical state thanks to our knowledge
of its preceding dynamics. In this example, though, both distributions are
usually immensely sharp and practically lead to the same predictions. In
the neuroscientific applications considered in this note the difference in
predictions of the sample \vs\ full-population applications may be very
relevant instead.
